---
title: "eth_new"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

```{r}
library(sf)
library(ggplot2)
library(readr)
library(dplyr)
library(tidyr)
```


```{r}
# auxiliary function for handling sums of NA vectors
# handles NA by returning NA if all are NA
sumNA <- function(x) {
  if (all(is.na(x))) {
    return(NA)
  } else {
    return(sum(x, na.rm=TRUE))
  }
}

# take a vector of inputs: if at least 1 input is 1, then return 1 else 0.
# handles NA by returning NA if all are NA
indicator <- function(x) {
  if (all(is.na(x))) {
    return(NA)
  } else if (sum(x, na.rm=TRUE) > 0) {
    return(1)
  } else {
    return(0)
  }
}
```



```{r}
eth_zone <- st_read('../shapefiles/eth/Eth_Zone_2013.shp')
eth_zone_sf <- st_transform(eth_zone, "+proj=longlat +datum=WGS84")
```


```{r}
# sect3 is individual level fertilizer use
# remark: read_csv from readr is apparently faster than base read.csv
sect3_pp_w1 <- read_csv('../lsms_data/ethiopia/sect3_pp_w1.csv')
sect3_pp_w2 <- read_csv('../lsms_data/ethiopia/sect3_pp_w2.csv') 
sect3_pp_w3 <- read_csv('../lsms_data/ethiopia/sect3_pp_w3.csv')

#sect4 is individual level herb/pest/fungicide use
sect4_pp_w1 <- read_csv('../lsms_data/ethiopia/sect4_pp_w1.csv')
sect4_pp_w2 <- read_csv('../lsms_data/ethiopia/sect4_pp_w2.csv')
sect4_pp_w3 <- read_csv('../lsms_data/ethiopia/sect4_pp_w3.csv')

# sect7 is household level (chemical fertilizer y/n). UNUSED
#sect7_pp_w1 <- read_csv('../lsms_data/ethiopia/sect7_pp_w1.csv')
#sect7_pp_w2 <- read_csv('../lsms_data/ethiopia/sect7_pp_w2.csv')
#sect7_pp_w3 <- read_csv('../lsms_data/ethiopia/sect7_pp_w3.csv')
```



# Plan

We'll do fertilizer separately from Pest/herb/fung data. Since data formats are almost same across all waves, will do the following:

1) join each wave's data with wave's location data
2) apply relevant transformations, then do sf_join to assign a zone geometry object (we'll call it SF_ZONE) to each row
3) group by the SF_ZONE name
4) merge with GAHI by SF_ZONE

# fertilizer data


```{r}
# get location data (Lon/Lat), and need to join later tables by ea_id.
geo1 <- read_csv('../lsms_data/ethiopia/pub_eth_householdgeovariables_y1.csv')
loc1 <- distinct(geo1 %>% select(ea_id, longitude=LON_DD_MOD, latitude=LAT_DD_MOD))

geo2 <- read_csv('../lsms_data/ethiopia/Pub_ETH_HouseholdGeovars_Y2.csv')
loc2 <- distinct(geo2 %>% select(ea_id, longitude=lon_dd_mod, latitude=lat_dd_mod))


#  we'll use household_id here as the primary key
geo3 <- read_csv('../lsms_data/ethiopia/ETH_HouseholdGeovars_y3.csv')
loc3 <- distinct(geo3 %>% select(household_id2, longitude=lon_dd_mod, latitude=lat_dd_mod))

```



```{r}

wave1 <- sect3_pp_w1 %>% select(ea_id, household_id, region=saq01, zone=saq02, fert=pp_s3q14, urea=pp_s3q15, dap=pp_s3q18, manure=pp_s3q21, compost=pp_s3q23, organic_fert=pp_s3q25) %>% drop_na(fert, household_id)

wave2 <- sect3_pp_w2 %>% select(ea_id, household_id, region=saq01, zone=saq02, fert=pp_s3q14, urea=pp_s3q15, dap=pp_s3q18, manure=pp_s3q21, compost=pp_s3q23, organic_fert=pp_s3q25) %>% drop_na(fert, household_id)

# wave 3 has location data by ea_id2, so we need to use that instead
wave3 <- sect3_pp_w3 %>% select(household_id, household_id2, region=saq01, zone=saq02, fert=pp_s3q14, urea=pp_s3q15, dap=pp_s3q18, manure=pp_s3q21, compost=pp_s3q23, organic_fert=pp_s3q25) %>% drop_na(fert, household_id)

```


```{r}

wave1 <- inner_join(loc1, wave1) %>% select(-ea_id)
wave1

wave2 <- inner_join(loc2, wave2) %>% select(-ea_id)
wave2

wave3 <- inner_join(loc3, wave3) %>% select(-household_id2)
wave3

```



```{r}
# add countryname/year data, and combine into one frame
w1_fert <- data.frame(country='ethiopia', start_yr = 2011, end_yr = 2012, wave1, stringsAsFactors = F) 
w2_fert <- data.frame(country='ethiopia', start_yr = 2013, end_yr = 2014, wave2, stringsAsFactors = F)
w3_fert <- data.frame(country='ethiopia', start_yr = 2015, end_yr = 2016, wave3, stringsAsFactors = F)
fert <- rbind(w1_fert, w2_fert, w3_fert)
fert
```

```{r}
# handle missing long/lat data - only 8 cases
fert <- fert[complete.cases(fert$longitude, fert$latitude), ]
```

```{r}
# spatial join with location data - assigns each lon/lat pair to a zone 
fert_sf <- st_as_sf(x=fert, coords=c('longitude', 'latitude'), crs= "+proj=longlat +datum=WGS84")

fert_joined  <- st_join(fert_sf, eth_zone_sf, left=FALSE)
fert_joined <- data.frame(fert_joined %>% select(1:13), stringsAsFactors = F)
```

```{r}
fert_joined
```

```{r}
# reordering
fert_joined <- fert_joined %>% select(country, start_yr, end_yr, household_id, region, zone, ZONENAME, fert, urea, dap, manure, compost, organic_fert)
fert_joined <- fert_joined %>% mutate(ZONENAME=as.character(ZONENAME))
```

# applying transformations

```{r}


# 1 encoded as YES
# 2 encoded as NO - wish to change to 0
# note: this notation allows us to modify anything except cols 1 to 7
fert_joined[, -c(1:7)][fert_joined[, -c(1:7)] == 2] <- 0

# handling NAs - assume that if fert is 0 then no fertilizer was used, so set other columns to 0
fert_joined[fert_joined$fert==0, 8:ncol(fert_joined)] <- 0


# summarize fertilizer use into 3 dummy vars: organic, inorganic, overall
fert_joined <- fert_joined %>%
  mutate(
    organic_fert_dummy = as.numeric((manure+compost+organic_fert) > 0),
    inorganic_fert_dummy = as.numeric((urea+dap) > 0),
    fert_dummy = as.numeric((fert+urea+dap+manure+compost>0)))


# compute number of uses per region|unique household
# see above function for indicator. final NAs are handled here.
# for a column, if all household observations are all NA then drop
fert_by_hh <- fert_joined %>% select(-c(region, zone)) %>% group_by(country, start_yr, end_yr, ZONENAME, household_id) %>%
  summarize_all(indicator) %>% drop_na()
# compute zone averages

fert_avg <- fert_by_hh %>% select(-household_id) %>% group_by(country, start_yr, end_yr,  ZONENAME) %>% summarize_all(mean)

# compute zone counts


fert_avg$num_hh_fert <- (fert_by_hh %>% tally())$n
fert_avg

```



# GAHI prevalence data
```{r}
africa <- read_csv('../GAHI/data/africa_data.csv')
eth <- africa %>% filter(country_title == 'Ethiopia') %>% select(publication_year, prevalence, number_examined, number_positive, longitude, latitude) %>% drop_na()
eth_sf <- st_as_sf(x=eth, coords=c('longitude', 'latitude'), crs= "+proj=longlat +datum=WGS84")
#plot(st_geometry(sp), axes=TRUE)
#plot(africa_pt, pch=1, col='red', add=TRUE)
joined_gahi <- st_join(eth_zone_sf, eth_sf, left=FALSE)
```

```{r}
joined_gahi
```


```{r}
joined_gahi <- joined_gahi %>% select(publication_year, number_examined, number_positive, ZONENAME) %>% group_by(publication_year, ZONENAME) %>% summarize(prevalence=sum(number_positive)/sum(number_examined), num_examined=sum(number_examined), num_positive=sum(number_positive))
joined_gahi
```

remark: maybe do the pesticide data first, combine fert + pest, then combine with GAHI

```{r}
fert_avg<- inner_join(joined_gahi, fert_avg, by = c('publication_year' = 'start_yr', 'ZONENAME' = 'ZONENAME'))
```

```{r}
fert_avg
```





# pesticide/herbicide/fungicide data

in retrospect, could have merged location data using household_id, but same result

```{r}
w1_phf <- sect4_pp_w1 %>% select(ea_id, household_id, region=saq01, zone=saq02, prevention=pp_s4q04, pest=pp_s4q05, herb=pp_s4q06, fung=pp_s4q07) %>% drop_na(household_id, prevention) 

w2_phf <- sect4_pp_w2 %>% select(ea_id, household_id, region=saq01, zone=saq02, prevention=pp_s4q04, pest=pp_s4q05, herb=pp_s4q06, fung=pp_s4q07) %>% drop_na(household_id, prevention) 

phf <- sect4_pp_w3 %>% select(household_id2, household_id, region=saq01, zone=saq02, prevention=pp_s4q04, pest=pp_s4q05, herb=pp_s4q06, fung=pp_s4q07) %>% drop_na(household_id, prevention) 


wave1 <- inner_join(loc1, w1_phf) %>% select(-ea_id)
wave1

wave2 <- inner_join(loc2, w2_phf) %>% select(-ea_id)
wave2

wave3 <- inner_join(loc3, phf) %>% select(-household_id2)
wave3
```

```{r}
w1_phf <- data.frame(country='ethiopia', start_yr = 2011, end_yr = 2012, wave1, stringsAsFactors = F) 
w2_phf <- data.frame(country='ethiopia', start_yr = 2013, end_yr = 2014, wave2, stringsAsFactors = F)
phf <- data.frame(country='ethiopia', start_yr = 2015, end_yr = 2016, wave3, stringsAsFactors = F)
phf <- rbind(w1_phf, w2_phf, phf)
phf
```

```{r}
# handle missing long/lat data - only 4 cases
phf <- phf[complete.cases(phf$longitude, phf$latitude), ]
# spatial join with location data - assigns each lon/lat pair to a zone 
phf_sf <- st_as_sf(x=phf, coords=c('longitude', 'latitude'), crs= "+proj=longlat +datum=WGS84")

phf <- st_join(phf_sf, eth_zone_sf, left=FALSE)
phf <- data.frame(phf %>% select(1:11), stringsAsFactors = F)
phf
```




```{r}
# reordering
phf <- phf %>% select(country, start_yr, end_yr, household_id, region, zone, ZONENAME, prevention, pest, herb, fung)
phf <- phf %>% mutate(ZONENAME=as.character(ZONENAME))

phf[, -c(1:7)][phf[, -c(1:7)]==2] <- 0

phf[phf$prevention==0, 8:ncol(phf)] <- 0

phf <- phf %>% group_by(country, start_yr, end_yr, ZONENAME, household_id) %>% 
  summarize_all(indicator) %>% drop_na()

# compute zone avgs
phf_avg <- phf %>% select(-household_id) %>% group_by(country, start_yr, end_yr, ZONENAME) %>% summarize_all(mean)

# compute zone counts
phf_avg$num_hh_phf <- (phf %>% tally())$n
phf_avg
```

```{r}
eth_final <- inner_join(fert_avg, phf_avg)
eth_final
```


```{r}
eth_final <- eth_final %>% select(country, ZONENAME, gahi_pub_yr=publication_year, lsms_start_yr=start_yr, lsms_end_yr=end_yr, prevalence, num_positive, num_examined, fert, urea, dap, manure, compost, organic_fert, organic_fert_dummy, inorganic_fert_dummy, fert_dummy, num_hh_fert, prevention, pest, herb, fung, num_hh_phf)
eth_final
```

```{r, eval=FALSE}
write_csv(eth_final, '../results/eth_final.csv')
```

