---
title: "eth_new"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)
```

```{r}
library(sf)
library(ggplot2)
library(readr)
library(dplyr)
library(tidyr)
```


```{r}
# auxiliary function for handling sums of NA vectors
# handles NA by returning NA if all are NA
sumNA <- function(x) {
  if (all(is.na(x))) {
    return(NA)
  } else {
    return(sum(x, na.rm=TRUE))
  }
}

# take a vector of inputs: if at least 1 input is 1, then return 1 else 0.
# handles NA by returning NA if all are NA
indicator <- function(x) {
  if (all(is.na(x))) {
    return(NA)
  } else if (sum(x, na.rm=TRUE) > 0) {
    return(1)
  } else {
    return(0)
  }
}
```



```{r}
eth_zone <- st_read("../shapefiles/eth/GADM/gadm36_ETH_2.shp")
eth_zone_sf <- st_transform(eth_zone, "+proj=longlat +datum=WGS84 +nodefs")
unique(eth_zone$NAME_2)
```


```{r}
# sect3 is individual level fertilizer use
# remark: read_csv from readr is apparently faster than base read.csv
sect3_pp_w1 <- read_csv("../lsms_data/ethiopia/sect3_pp_w1.csv")
sect3_pp_w2 <- read_csv("../lsms_data/ethiopia/sect3_pp_w2.csv") 
sect3_pp_w3 <- read_csv("../lsms_data/ethiopia/sect3_pp_w3.csv")

#sect4 is individual level herb/pest/fungicide use
sect4_pp_w1 <- read_csv("../lsms_data/ethiopia/sect4_pp_w1.csv")
sect4_pp_w2 <- read_csv("../lsms_data/ethiopia/sect4_pp_w2.csv")
sect4_pp_w3 <- read_csv("../lsms_data/ethiopia/sect4_pp_w3.csv")

# sect7 is household level (chemical fertilizer y/n). UNUSED
#sect7_pp_w1 <- read_csv("../lsms_data/ethiopia/sect7_pp_w1.csv")
#sect7_pp_w2 <- read_csv("../lsms_data/ethiopia/sect7_pp_w2.csv")
#sect7_pp_w3 <- read_csv("../lsms_data/ethiopia/sect7_pp_w3.csv")
```



# Plan

We"ll do fertilizer separately from Pest/herb/fung data. Since data formats are almost same across all waves, will do the following:

1) join each wave"s data with wave"s location data
2) apply relevant transformations, then do sf_join to assign a zone geometry object & NAME_2 to each row
3) group by NAME_2
4) merge GAHI + LSMS by NAME_2 and year. 

# fertilizer data


```{r}
# get location data (Lon/Lat), and need to join later tables by ea_id.
geo1 <- read_csv("../lsms_data/ethiopia/pub_eth_householdgeovariables_y1.csv")
loc1 <- geo1 %>% select(household_id, longitude=LON_DD_MOD, latitude=LAT_DD_MOD)

geo2 <- read_csv("../lsms_data/ethiopia/Pub_ETH_HouseholdGeovars_Y2.csv")
loc2 <- geo2 %>% select(household_id2, longitude=lon_dd_mod, latitude=lat_dd_mod)


#   the fertilizer file is missing ea_id2, so we need tot use household_id2 to join here. 
geo3 <- read_csv("../lsms_data/ethiopia/ETH_HouseholdGeovars_y3.csv")
loc3 <- geo3 %>% select(household_id2, longitude=lon_dd_mod, latitude=lat_dd_mod)

loc1
loc2
loc3
```



```{r}

wave1 <- sect3_pp_w1 %>% select(household_id, region=saq01, zone=saq02, fert=pp_s3q14, urea=pp_s3q15, dap=pp_s3q18, manure=pp_s3q21, compost=pp_s3q23, organic_fert=pp_s3q25, quant_urea=pp_s3q16_a, quant_dap=pp_s3q19_a, area=pp_s3q05_a) %>% drop_na(fert, household_id)

wave2 <- sect3_pp_w2 %>% select(household_id2, region=saq01, zone=saq02, fert=pp_s3q14, urea=pp_s3q15, dap=pp_s3q18, manure=pp_s3q21, compost=pp_s3q23, organic_fert=pp_s3q25, quant_urea=pp_s3q16_a, quant_dap=pp_s3q19_a, area=pp_s3q05_a) %>% drop_na(fert, household_id2)

# wave 3 has location data by ea_id2, so we need to use household_id2 to match with the ea_id2 to finally get the location data
wave3 <- sect3_pp_w3 %>% select(household_id2, region=saq01, zone=saq02, fert=pp_s3q14, urea=pp_s3q15, dap=pp_s3q18, manure=pp_s3q21, compost=pp_s3q23, organic_fert=pp_s3q25, quant_urea=pp_s3q16, quant_dap=pp_s3q19, area=pp_s3q05_a) %>% drop_na(fert, household_id2)

wave1
wave2
wave3
```


Note on units: 
field area: square meters
urea quant: kg (rounded down)
dap quant: kg (rounded down)

so quantity/area is in kg/m^2




```{r}

wave1_j <- inner_join(wave1, loc1) 
wave1_j

wave2_j <- inner_join(wave2, loc2) %>% rename(household_id = household_id2)
wave2_j

wave3_j <- inner_join(wave3, loc3) %>% rename(household_id = household_id2)
wave3_j

```



```{r}
# add countryname/year data, and combine into one frame
w1_fert <- data.frame(country="ethiopia", start_yr = 2011, end_yr = 2012, wave1_j, stringsAsFactors = F) 
w2_fert <- data.frame(country="ethiopia", start_yr = 2013, end_yr = 2014, wave2_j, stringsAsFactors = F)
w3_fert <- data.frame(country="ethiopia", start_yr = 2015, end_yr = 2016, wave3_j, stringsAsFactors = F)
fert <- rbind(w1_fert, w2_fert, w3_fert)
fert
```

```{r}
# handle missing long/lat data - only 8 cases
fert <- fert[complete.cases(fert$longitude, fert$latitude), ]
```

```{r}
# spatial join with location data - assigns each lon/lat pair to a zone 
fert_sf <- st_as_sf(x=fert, coords=c("longitude", "latitude"), crs= "+proj=longlat +datum=WGS84 +nodefs")

fert_joined  <- st_join(fert_sf, eth_zone_sf, left=FALSE) %>% select(c(1:15), NAME_2)
```

```{r}
fert_df <- fert_joined %>% st_set_geometry(NULL)
fert_df
```

```{r}
# reordering
fert_only <- fert_df %>% select(country, start_yr, end_yr, household_id, NAME_2, fert, urea, dap, manure, compost, organic_fert) %>% mutate(NAME_2=as.character(NAME_2))
fert_only

# quantity data
fert_quant <- fert_df %>% select(country, start_yr, end_yr, household_id, NAME_2, quant_urea, quant_dap, area) %>% filter(complete.cases(.))
fert_quant
```
Remark: we have about 70k observations for fert data, but only about 7k observations for quantity/area. When we group by household, this number will decrease further.

# applying transformations to fert_only

```{r}

# 1 encoded as YES
# 2 encoded as NO - wish to change to 0
recode <- function(x) ifelse(x == 2 | x == 0, 0, 1)
fert_only <- fert_only %>% mutate_at(c("fert", "urea", "dap", "manure", "compost", "organic_fert"), recode)

# handling NAs - assume that if fert is 0 then no fertilizer was used, so set other columns to 0
# remark: dot in list is essentially a placeholder for all variables that we want to pass to it
# remark2: funs is deprecated in favor of list in dplyr 0.8.0
handleNA <- function(x, y) ifelse(x == 0, 0, y)
fert_only <- fert_only %>% mutate_at(c("urea", "dap", "manure", "compost", "organic_fert"), list(~handleNA(fert, .)) )


# summarize fertilizer use into 3 dummy vars: organic, inorganic, overall
fert_only <- fert_only %>%
  mutate(
    organic_fert_dummy = as.numeric((manure+compost+organic_fert) > 0),
    inorganic_fert_dummy = as.numeric((urea+dap) > 0),
    fert_dummy = as.numeric((fert+urea+dap+manure+compost>0)))

# compute number of uses per zone|unique household
# see above function for indicator. final NAs are handled here.
# for each household: if all obs are NA then return NA and drop, else return 1 if atleast one member used fertilizer or 0
fert_by_hh <- fert_only %>% group_by(country, start_yr, end_yr, NAME_2, household_id) %>%
  summarize_all(indicator) %>% drop_na()

# compute zone averages
fert_avg <- fert_by_hh %>% select(-household_id) %>% group_by(country, start_yr, end_yr,  NAME_2) %>% summarize_all(mean)

# compute count of households/zone
fert_avg$num_hh_fert <- (fert_by_hh %>% tally())$n
fert_avg

```

# applying transformations to fert_quant
```{r}
# per zone: take total quantity/total area
fert_quant_avg <- fert_quant %>% group_by(country, start_yr, end_yr, NAME_2) %>% summarize(total_urea = sum(quant_urea), total_dap=sum(quant_dap), total_area = sum(area), urea_use = total_urea/total_area, dap_use = total_dap/total_area, combined_use = (total_urea+total_dap)/total_area) 
fert_quant_avg$num_hh_quant <- (fert_quant %>% group_by(country, start_yr, end_yr, NAME_2) %>% tally())$n
fert_quant_avg
```


```{r}
fert <- full_join(fert_avg, fert_quant_avg)
fert
```


# LSMS pesticide/herbicide/fungicide data

in retrospect, could have merged location data using household_id, but same result.

Note: no data on quantity used is availible.

```{r}
w1_phf <- sect4_pp_w1 %>% select( household_id, region=saq01, zone=saq02, prevention=pp_s4q04, pest=pp_s4q05, herb=pp_s4q06, fung=pp_s4q07) %>% drop_na(household_id, prevention) 

w2_phf <- sect4_pp_w2 %>% select(household_id2, region=saq01, zone=saq02, prevention=pp_s4q04, pest=pp_s4q05, herb=pp_s4q06, fung=pp_s4q07) %>% drop_na(household_id2, prevention) 

w3_phf <- sect4_pp_w3 %>% select(household_id2, region=saq01, zone=saq02, prevention=pp_s4q04, pest=pp_s4q05, herb=pp_s4q06, fung=pp_s4q07) %>% drop_na(household_id2, prevention) 


wave1 <- inner_join(w1_phf, loc1) 
wave1

wave2 <- inner_join(w2_phf, loc2) %>% rename(household_id = household_id2)
wave2

wave3 <- inner_join(w3_phf, loc3) %>% rename(household_id = household_id2)
wave3
```

```{r}
w1_phf <- data.frame(country="ethiopia", start_yr = 2011, end_yr = 2012, wave1, stringsAsFactors = F) 
w2_phf <- data.frame(country="ethiopia", start_yr = 2013, end_yr = 2014, wave2, stringsAsFactors = F)
phf <- data.frame(country="ethiopia", start_yr = 2015, end_yr = 2016, wave3, stringsAsFactors = F)
phf <- rbind(w1_phf, w2_phf, phf)
phf
```

```{r}
# handle missing long/lat data - only 4 cases
phf <- phf[complete.cases(phf$longitude, phf$latitude), ]
# spatial join with location data - assigns each lon/lat pair to a zone 
phf_sf <- st_as_sf(x=phf, coords=c("longitude", "latitude"), crs= "+proj=longlat +datum=WGS84 +nodefs")

phf_joined <- st_join(phf_sf, eth_zone_sf, left=FALSE)
phf_df <- phf_joined %>% st_set_geometry(NULL) 
phf_df <- phf_df %>% mutate(NAME_2=as.character(NAME_2)) %>% select(country, start_yr, end_yr, NAME_2, household_id, prevention, pest, herb, fung) 
phf_df
```




```{r}
# 1 encoded as YES
# 2 encoded as NO - wish to change to 0
recode <- function(x) ifelse(x == 2, 0, 1)
phf_df <- phf_df %>% mutate_at(c("prevention", "pest", "herb", "fung"), recode)

# handling NAs - assume that if prevention is 0 then set other columns to 0
handleNA <- function(x, y) ifelse(x == 0, 0, y)
phf <- phf_df %>% mutate_at(c("pest", "herb", "fung"), list(~handleNA(prevention, .)) )


phf_df <- phf_df %>% group_by(country, start_yr, end_yr, NAME_2, household_id) %>% 
  summarize_all(indicator) %>% drop_na()

# compute zone avgs
phf_avg <- phf_df %>% select(-household_id) %>% group_by(country, start_yr, end_yr, NAME_2) %>% summarize_all(mean)

# compute zone counts
phf_avg$num_hh_phf <- (phf_df %>% tally())$n
phf_avg
```


```{r}
eth_lsms <- full_join(fert, phf_avg)
eth_lsms
```

# GAHI prevalence data
```{r}
africa <- read_csv("../GAHI/data/africa_data.csv")
eth <- africa %>% filter(country_title == "Ethiopia") %>% select(publication_year, prevalence, number_examined, number_positive, longitude, latitude) %>% drop_na()
eth_sf <- st_as_sf(x=eth, coords=c("longitude", "latitude"), crs= "+proj=longlat +datum=WGS84 +nodefs")
#plot(st_geometry(sp), axes=TRUE)
#plot(africa_pt, pch=1, col="red", add=TRUE)
joined_gahi <- st_join(eth_zone_sf, eth_sf, left=FALSE)
joined_gahi
```



```{r}
eth_gahi <- joined_gahi %>% select(publication_year, number_examined, number_positive, NAME_2) %>% group_by(publication_year, NAME_2) %>% summarize(prevalence=sum(number_positive)/sum(number_examined), num_examined=sum(number_examined), num_positive=sum(number_positive))
eth_gahi 
```



```{r}
eth_gahi_lsms<- inner_join(eth_gahi, eth_lsms, by = c("publication_year" = "start_yr", "NAME_2" = "NAME_2")) %>% st_set_geometry(NULL)
eth_gahi_lsms
```


# Final dataset 
To summarize, lsms data was outer joined together, and then lsms + gahi was inner joined.
```{r}
eth_final <- eth_gahi_lsms %>% select(country, level2_name=NAME_2, gahi_lsms_start_yr=publication_year, lsms_end_yr=end_yr, prevalence, num_positive, num_examined, fert, urea, dap, manure, compost, organic_fert, organic_fert_dummy, inorganic_fert_dummy, fert_dummy, num_hh_fert, total_urea, total_dap, total_area, urea_use, dap_use, num_hh_quant, prevention, pest, herb, fung, num_hh_phf)
eth_final
```


```{r, eval=FALSE}
#exporting: simply run to write file
write_csv(eth_final, "../results/eth_final.csv")
```


# Plots


```{r}
pairs(~prevalence + fert + urea + dap + pest + herb + fung, data=eth_final)
```

need more data!








