---
title: "Ethiopia LSMS-ISA"
author: "Elliot Quan"
date: "2/26/2019""
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```

```{r}
library(ggplot2)
library(readr)
library(dplyr)
library(tidyr)
#library(haven) # for converting .dta to .csv - not needed anymore
```



```{r}
# auxiliary function for handling sums of NA vectors
# handles NA by returning NA if all are NA
sumNA <- function(x) {
  if (all(is.na(x))) {
    return(NA)
  } else {
    return(sum(x, na.rm=TRUE))
  }
}

# take a vector of inputs: if at least 1 input is 1, then return 1 else 0.
# handles NA by returning NA if all are NA
indicator <- function(x) {
  if (all(is.na(x))) {
    return(NA)
  } else if (sum(x, na.rm=TRUE) > 0) {
    return(1)
  } else {
    return(0)
  }
}
```


```{r}
# sect3 is individual level fertilizer use
# remark: read_csv from readr is apparently faster than base read.csv
sect3_pp_w1 <- read_csv('../lsms_data/ethiopia/sect3_pp_w1.csv')
sect3_pp_w2 <- read_csv('../lsms_data/ethiopia/sect3_pp_w2.csv') 
sect3_pp_w3 <- read_csv('../lsms_data/ethiopia/sect3_pp_w3.csv')

#sect4 is individual level herb/pest/fungicide use
sect4_pp_w1 <- read_csv('../lsms_data/ethiopia/sect4_pp_w1.csv')
sect4_pp_w2 <- read_csv('../lsms_data/ethiopia/sect4_pp_w2.csv')
sect4_pp_w3 <- read_csv('../lsms_data/ethiopia/sect4_pp_w3.csv')

# sect7 is household level (chemical fertilizer y/n). UNUSED
#sect7_pp_w1 <- read_csv('../lsms_data/ethiopia/sect7_pp_w1.csv')
#sect7_pp_w2 <- read_csv('../lsms_data/ethiopia/sect7_pp_w2.csv')
#sect7_pp_w3 <- read_csv('../lsms_data/ethiopia/sect7_pp_w3.csv')
```




In final output, each row should correspond to country|year|region|zone, where region and zone are the 1st
and 2nd subdivisions of Ethiopia.

# Wave 1 (2011-2012)

This is the fert data (sect3_pp)
```{r} 
wave1 <- sect3_pp_w1 %>% 
  select(household_id, region=saq01, zone=saq02, fert=pp_s3q14, urea=pp_s3q15, dap=pp_s3q18, 
         manure=pp_s3q21, compost=pp_s3q23, organic_fert=pp_s3q25) %>% 
  drop_na(fert)

# wave1 region 12 has 0 fertilizer use.
unique(wave1[wave1$region == 12,]['fert']) 

# 1 encoded as YES
# 2 encoded as NO - wish to change to 0
# note: this notation allows us to modify anything except cols 1, 2, 3
wave1[, -c(1,2,3)][wave1[, -c(1,2,3)] == 2] <- 0

# handling NAs - assume that if fert is 0 then no fertilizer was used, so set other columns to 0
wave1[wave1$fert==0, 4:ncol(wave1)] <- 0


# summarize fertilizer use into 3 dummy vars: organic, inorganic, overall
wave1 <- wave1 %>%
  mutate(
    organic_fert_dummy = as.numeric((manure+compost+organic_fert) > 0),
    inorganic_fert_dummy = as.numeric((urea+dap) > 0),
    fert_dummy = as.numeric((fert+urea+dap+manure+compost>0)))


# compute number of uses per region|unique household
# see above function for indicator. final NAs are handled here.
# for a column, if all household observations are all NA then drop
w1_fert <- wave1 %>% group_by(region, zone, household_id) %>%
  summarize_all(indicator) %>% drop_na()


# compute zone averages
w1_fert_avg <- w1_fert %>% select(-household_id) %>% group_by(region, zone) %>% summarize_all(mean)

# compute zone counts
w1_fert_avg$num_hh_fert <- (w1_fert %>% tally())$n
w1_fert_avg
```


Now we compute the p/h/f data in the same manner.

```{r}
w1_phf <- sect4_pp_w1 %>% select(household_id, region=saq01, zone=saq02, prevention=pp_s4q04, pest=pp_s4q05, herb=pp_s4q06, fung=pp_s4q07) %>% drop_na(prevention) # if prevention method not specified then drop - can't recover.

# encode 2s (NO) to 0s for columns 4 and on 
w1_phf[, -c(1,2,3)][w1_phf[, -c(1,2,3)] == 2] <- 0


# if no prevention used, set all rest of columns to 0
w1_phf[w1_phf$prevention==0, 4:ncol(w1_phf)] <- 0

# group by household level
w1_phf <- w1_phf %>% group_by(region, zone, household_id) %>% summarize_all(indicator) %>% drop_na()

# compute zone avgs
w1_phf_avg <- w1_phf %>% select(-household_id) %>% group_by(region, zone) %>% summarize_all(mean)

# compute zone counts
w1_phf_avg$num_hh_phf <- (w1_phf %>% tally())$n
w1_phf_avg
```


```{r}
# zones match thankfully -  no missing zones.
w1_final <- w1_fert_avg %>% inner_join(w1_phf_avg)
w1_final
```


Something that stood out: region 12 has 0 reported fertilizer use.

```{r}
r12 <- sect3_pp_w1 %>% 
  select(household_id, region=saq01, zone=saq02, fert=pp_s3q14, urea=pp_s3q15, dap=pp_s3q18, 
         manure=pp_s3q21, compost=pp_s3q23, organic_fert=pp_s3q25) %>% 
  drop_na(fert)
unique(r12[r12$region == 12, ]$fert)
```

# Wave 2 (2013-2014)

Luckily structure of wave2 data is almost identical to wave 1, so we can copy/paste most of the code. It may be more prudent to write a function (which I may do later, depending on how similar other survey data is) to do all of this

```{r}
wave2 <- sect3_pp_w2 %>% 
  select(household_id, region=saq01, zone=saq02, fert=pp_s3q14, urea=pp_s3q15, dap=pp_s3q18, manure=pp_s3q21, compost=pp_s3q23, organic_fert=pp_s3q25) %>% 
  drop_na(fert)

wave2[, -c(1,2,3)][wave2[, -c(1,2,3)] == 2] <- 0

wave2[wave2$fert==0, 4:ncol(wave2)] <- 0


wave2 <- wave2 %>%
  mutate(
    organic_fert_dummy = as.numeric((manure+compost+organic_fert) > 0),
    inorganic_fert_dummy = as.numeric((urea+dap) > 0),
    fert_dummy = as.numeric((fert+urea+dap+manure+compost)>0))


w2_fert <- wave2 %>% group_by(region, zone, household_id) %>%
  summarize_all(indicator) %>% drop_na()

# compute zone averages
w2_fert_avg <- w2_fert %>% select(-household_id) %>% group_by(region, zone) %>% summarize_all(mean)

# compute zone counts
w2_fert_avg$num_hh_fert <- (w2_fert %>% tally())$n
w2_fert_avg
```


p/h/f data
```{r}
w2_phf <- sect4_pp_w2 %>% select(household_id, region=saq01, zone=saq02, prevention=pp_s4q04, pest=pp_s4q05, herb=pp_s4q06, fung=pp_s4q07) %>% drop_na(prevention) 

# set 2s to 0s
w2_phf[, -c(1,2,3)][w2_phf[, -c(1,2,3)] == 2] <- 0


# if no prevention used, set all rest of columns to 0
w2_phf[w2_phf$prevention==0, 4:ncol(w2_phf)] <- 0

w2_phf <- w2_phf %>% group_by(region, zone, household_id) %>% summarize_all(indicator) %>% drop_na()


# compute zone avgs
w2_phf_avg <- w2_phf %>% select(-household_id) %>% group_by(region, zone) %>% summarize_all(mean)

# compute zone counts
w2_phf_avg$num_hh_phf <- (w2_phf %>% tally())$n
w2_phf_avg
```

```{r}
# remark: there are missing zones in w2_phf, so NA rows are kept in the left_join. to remove missing zones
# just use inner join (or add drop_na())
w2_final <- w2_fert_avg %>% left_join(w2_phf_avg)
w2_final
```


# Wave 3 (2015-2016)


```{r}
wave3 <- sect3_pp_w3 %>% 
  select(household_id, region=saq01, zone=saq02, fert=pp_s3q14, urea=pp_s3q15, dap=pp_s3q18, manure=pp_s3q21, compost=pp_s3q23, organic_fert=pp_s3q25) %>% 
  drop_na(fert)

wave3[, -c(1,2,3)][wave3[, -c(1,2,3)] == 2] <- 0

wave3[wave3$fert==0, 4:ncol(wave3)] <- 0

wave3 <- wave3 %>%
  mutate(
    organic_fert_dummy = as.numeric((manure+compost+organic_fert) > 0),
    inorganic_fert_dummy = as.numeric((urea+dap) > 0),
    fert_dummy = as.numeric((fert+urea+dap+manure+compost)>0)) %>% drop_na()

# compute number of uses per region|unique household_
w3_fert <- wave3 %>% group_by(region, zone, household_id) %>% 
  summarize_all(indicator) %>% drop_na()

# compute zone averages
w3_fert_avg <- w3_fert %>% select(-household_id) %>% group_by(region, zone) %>% summarize_all(mean)

# compute zone counts
w3_fert_avg$num_hh_fert <- (w3_fert %>% tally())$n
w3_fert_avg
```


p/h/f data

```{r}
w3_phf <- sect4_pp_w3 %>% select(household_id, region=saq01, zone=saq02, prevention=pp_s4q04, pest=pp_s4q05, herb=pp_s4q06, fung=pp_s4q07) %>% drop_na(prevention) 

w3_phf[, -c(1,2,3)][w3_phf[, -c(1,2,3)]==2] <- 0

w3_phf[w3_phf$prevention==0, 4:ncol(w3_phf)] <- 0

w3_phf <- w3_phf %>% group_by(region, zone, household_id) %>% 
  summarize_all(indicator) %>% drop_na()

# compute zone avgs
w3_phf_avg <- w3_phf %>% select(-household_id) %>% group_by(region, zone) %>% summarize_all(mean)

# compute zone counts
w3_phf_avg$num_hh_phf <- (w3_phf %>% tally())$n
w3_phf_avg
```

```{r}
w3_final <- w3_fert_avg %>% inner_join(w3_phf_avg)
```




# Combined Data

```{r}
eth_w1 <- data.frame(country=rep('ethiopia', nrow(w1_final)), start_yr = 2011, end_yr = 2012, w1_final)
eth_w2 <- data.frame(country=rep('ethiopia', nrow(w2_final)), start_yr = 2013, end_yr = 2014, w2_final)
eth_w3 <- data.frame(country=rep('ethiopia', nrow(w3_final)), start_yr = 2015, end_yr = 2016, w3_final)
```

```{r}
eth_final <- rbind(eth_w1, eth_w2, eth_w3) %>% rename(avg_rep_fert=fert, avg_urea=urea, avg_dap=dap, avg_manure = manure,
                                                            avg_compost=compost, avg_rep_ofert=organic_fert, 
                                                            avg_ofert = organic_fert_dummy,
                                                            avg_iofert = inorganic_fert_dummy,
                                                            avg_fert = fert_dummy, avg_prevention = prevention,
                                                            avg_pest=pest, avg_herb=herb, avg_fung=fung)
```


```{r, eval=FALSE}
write_csv(eth_final, path='../results/eth_lsms.csv')
```

```{r}
eth_final
```





# s7 wave1 (chem. fert.) data - UNUSED

```{r, eval=FALSE}
s7 <- sect7_pp_w1 %>%
  select(household_id, pw, chem_fert=pp_s7q02)
s7[s7==2] <- 0
s7_hh_w1 <- s7 %>% group_by(household_id, pw) %>% summarize(chem_fert=sum(chem_fert, na.rm=TRUE) > 0)
s7_hh_w1$chem_fert <- as.integer(s7_hh_w1$chem_fert)
mean(s7_hh_w1$chem_fert) # we see that without weighting, avg is 41%
s7_hh_w1 <- s7_hh_w1 %>% mutate(chem_fert_w=pw*chem_fert) 
sum(s7_hh_w1$chem_fert_w)/sum(s7_hh_w1$pw) # with weighting, avg is 54%
```






