---
title: 'Ethiopia LSMS + GAHI processing'
author: Elliot Quan
output: html_document
date: 3/29/19
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)
```

```{r}
library(sf)
library(tidyverse)
```

# Rough Plan

We"ll do fertilizer data separately from pest/herb/fung data, and then combine down the line. Since data formats are almost same across all waves, will do the following:

1) join each wave's data with wave's location data
2) apply relevant transformations, then do sf_join to assign a zone geometry object & NAME_2 (names of the second subdivisions of the country) to each row. 
3) group by household, applying indicators where necessary
3) group by NAME_2 
4) merge GAHI + LSMS by NAME_2 and year. 



# Custom Functions
```{r}
# auxiliary function for handling sums of NA vectors
# handles NA by returning NA if all are NA
sumNA <- function(x) {
  if (all(is.na(x))) {
    return(NA)
  } else {
    return(sum(x, na.rm=TRUE))
  }
}

# take a vector of inputs: if at least 1 input is 1, then return 1 else 0.
# handles NA by returning NA if all are NA
indicator <- function(x) {
  if (all(is.na(x))) {
    return(NA)
  } else if (sum(x, na.rm=TRUE) > 0) {
    return(1)
  } else {
    return(0)
  }
}
```


# Shapefile Data (GADM)
Note: there are two zones with the same name: North Shewa. 
```{r}
eth_zone <- st_read("../shapefiles/eth/GADM/gadm36_ETH_2.shp")
eth_zone_sf <- st_transform(eth_zone, crs=3857) # convert to mercator/planar (3857)
```

```{r}
eth_zone_sf[eth_zone_sf$NAME_1 == 'Amhara' & eth_zone_sf$NAME_2 == 'North Shewa', 'NAME_2'] 

```


# LSMS Agricultural Data
```{r}
# sect3 is individual level fertilizer use
# remark: read_csv from readr is apparently faster than base read.csv
sect3_pp_w1 <- read_csv("../lsms_data/ethiopia/sect3_pp_w1.csv")
sect3_pp_w2 <- read_csv("../lsms_data/ethiopia/sect3_pp_w2.csv") 
sect3_pp_w3 <- read_csv("../lsms_data/ethiopia/sect3_pp_w3.csv")

#sect4 is individual level herb/pest/fungicide use
sect4_pp_w1 <- read_csv("../lsms_data/ethiopia/sect4_pp_w1.csv")
sect4_pp_w2 <- read_csv("../lsms_data/ethiopia/sect4_pp_w2.csv")
sect4_pp_w3 <- read_csv("../lsms_data/ethiopia/sect4_pp_w3.csv")

# sect7 is household level (chemical fertilizer y/n). UNUSED, since redundant by sect4
#sect7_pp_w1 <- read_csv("../lsms_data/ethiopia/sect7_pp_w1.csv")
#sect7_pp_w2 <- read_csv("../lsms_data/ethiopia/sect7_pp_w2.csv")
#sect7_pp_w3 <- read_csv("../lsms_data/ethiopia/sect7_pp_w3.csv")
```



# LSMS geovariables (location data)

```{r}
# get location data (Lon/Lat) for each household_id
geo1 <- read_csv("../lsms_data/ethiopia/pub_eth_householdgeovariables_y1.csv")
loc1 <- geo1 %>% select(household_id, longitude=LON_DD_MOD, latitude=LAT_DD_MOD)

# for wave2/3, we use household_id2 (household_id applicable if household was sampled from wave1)
geo2 <- read_csv("../lsms_data/ethiopia/Pub_ETH_HouseholdGeovars_Y2.csv")
loc2 <- geo2 %>% select(household_id2, longitude=lon_dd_mod, latitude=lat_dd_mod)

geo3 <- read_csv("../lsms_data/ethiopia/ETH_HouseholdGeovars_y3.csv")
loc3 <- geo3 %>% select(household_id2, longitude=lon_dd_mod, latitude=lat_dd_mod)

loc1
loc2
loc3
```



```{r}
# selecting relevant columns 
wave1 <- sect3_pp_w1 %>% select(household_id, region=saq01, zone=saq02, fert=pp_s3q14, urea=pp_s3q15, dap=pp_s3q18, manure=pp_s3q21, compost=pp_s3q23, organic_fert=pp_s3q25, quant_urea=pp_s3q16_a, quant_dap=pp_s3q19_a, area=pp_s3q05_a) %>% drop_na(fert, household_id)

wave2 <- sect3_pp_w2 %>% select(household_id2, region=saq01, zone=saq02, fert=pp_s3q14, urea=pp_s3q15, dap=pp_s3q18, manure=pp_s3q21, compost=pp_s3q23, organic_fert=pp_s3q25, quant_urea=pp_s3q16_a, quant_dap=pp_s3q19_a, area=pp_s3q05_a) %>% drop_na(fert, household_id2)

wave3 <- sect3_pp_w3 %>% select(household_id2, region=saq01, zone=saq02, fert=pp_s3q14, urea=pp_s3q15, dap=pp_s3q18, manure=pp_s3q21, compost=pp_s3q23, organic_fert=pp_s3q25, quant_urea=pp_s3q16, quant_dap=pp_s3q19, area=pp_s3q05_a) %>% drop_na(fert, household_id2)

wave1
wave2
wave3
```


Note on units: 
field area: square meters
urea quant: kg (rounded down)
dap quant: kg (rounded down)

so quantity/area is in kg/m^2




```{r}
# joining fertilizer data with locations (lat/lon)
wave1_j <- inner_join(wave1, loc1) 
wave1_j

wave2_j <- inner_join(wave2, loc2) %>% rename(household_id = household_id2)
wave2_j

wave3_j <- inner_join(wave3, loc3) %>% rename(household_id = household_id2)
wave3_j

```



```{r}
# add countryname/year columns, and combine into one frame
w1_fert <- data.frame(country="ethiopia", start_yr = 2011, end_yr = 2012, wave1_j, stringsAsFactors = F) 
w2_fert <- data.frame(country="ethiopia", start_yr = 2013, end_yr = 2014, wave2_j, stringsAsFactors = F)
w3_fert <- data.frame(country="ethiopia", start_yr = 2015, end_yr = 2016, wave3_j, stringsAsFactors = F)
fert <- rbind(w1_fert, w2_fert, w3_fert)
fert
```

```{r}
# handle missing lon/lat data 
fert <- fert[complete.cases(fert$longitude, fert$latitude), ]
# store lon/lat cols for later
f_lon <- fert$longitude 
f_lat <- fert$latitude
```

```{r}
# spatial join with location data - assigns each lon/lat pair to a zone 
fert_sf <- st_as_sf(x=fert, coords=c("longitude", "latitude"), crs=4326)
fert_sf <- st_transform(fert_sf, crs=3857)
fert_sf

fert_joined  <- st_join(fert_sf, eth_zone_sf, left=FALSE) %>% select(c(1:15), NAME_1, NAME_2)
fert_df <- fert_joined %>% st_set_geometry(NULL)
fert_df
```

```{r}
# reordering columns
fert_only <- fert_df %>% select(country, start_yr, end_yr, household_id, NAME_1, NAME_2, fert, urea, dap, manure, compost, organic_fert) %>% mutate(NAME_1=as.character(NAME_1), NAME_2=as.character(NAME_2))
fert_only
```

# Fertilizer Quantity Data
```{r}
fert_quant <- fert_df %>% select(country, start_yr, end_yr, household_id, NAME_1, NAME_2, quant_urea, quant_dap, area) %>% mutate(total_quant = quant_urea + quant_dap) %>% drop_na()
```

We have some obvious outliers - these might warrant removal later.
```{r}
summary(fert_quant$area)
summary(fert_quant$quant_urea)
summary(fert_quant$quant_dap)
summary(fert_quant$total_quant)
```
Remark: we have about 70k observations for fert data, but only about 7k observations for quantity/area. When we group by household, this number will decrease further.

```{r}
plot(fert_quant$quant_urea)
```

Not sure what to do with these outliers.
```{r}
fert_quant[fert_quant$quant_urea > 2000, ]
```
Indeed we can see what happens when we don't remove extreme values - so let's remove them? seems highly unlikely that one person used 2500 kg of fertilizer...

# Cleaning/Transforming Fertilizer Data

```{r}
# LSMS data has 1 encoded as YES, 2 encoded as NO - we will change all 2s to 0s
recode <- function(x) ifelse(x == 2 | x == 0, 0, 1)
fert_only <- fert_only %>% mutate_at(c("fert", "urea", "dap", "manure", "compost", "organic_fert"), recode)

# handling NAs - assume that if fert is 0 then no fertilizer was used, so set other columns to 0
# remark 1: dot in list is essentially a placeholder for all variables that we want to pass to it
# remark 2: funs is deprecated in favor of list in dplyr 0.8.0
handleNA <- function(x, y) ifelse(x == 0, 0, y)
fert_only <- fert_only %>% mutate_at(c("urea", "dap", "manure", "compost", "organic_fert"), list(~handleNA(fert, .)) )


# summarize fertilizer use into 3 dummy vars: organic, inorganic, overall
# note that there is a fertilizer column from the LSMS data, but we're creating our own just in case the
# data is unreliable
fert_only <- fert_only %>%
  mutate(organic_fert_dummy = as.numeric((manure+compost+organic_fert) > 0),
    inorganic_fert_dummy = as.numeric((urea+dap) > 0),
    fert_dummy = as.numeric((fert+urea+dap+manure+compost>0)))

# compute number of uses per unique household using custom indicator function.
# final NAs are handled here.
fert_by_hh <- fert_only %>% group_by(country, start_yr, end_yr, NAME_1, NAME_2, household_id) %>%
  summarize_all(indicator) %>% drop_na()

# compute zone averages
fert_avg <- fert_by_hh %>% select(-household_id) %>% group_by(country, start_yr, end_yr,  NAME_1, NAME_2) %>% summarize_all(mean)

# compute count of households/zone
fert_avg$num_hh_fert <- (fert_by_hh %>% tally())$n
fert_avg
```

# Applying Transformations to Fert Quantity Data
```{r}
# we want the average household usage, so we'll just compute an average for each household, and then compute an average over households.
fert_quant_hh <- fert_quant %>% group_by(country, start_yr, end_yr, NAME_1, NAME_2, household_id) %>% 
  summarize(urea_use = mean(quant_urea/area), dap_use = mean(quant_dap/area), iofert_use = mean((quant_urea + quant_dap)/area))

# compute the mean usage of each zone (over households)
fert_quant_avg <- fert_quant_hh %>% select(-household_id) %>% group_by(country, start_yr, end_yr, NAME_1, NAME_2) %>% summarize_all(mean)

# compute the number of households per zone 
fert_quant_avg$num_hh_quant <- (fert_quant_hh %>% tally())$n
fert_quant_avg
```


```{r}
# full join used in order to preserve data from both sets - however we should expect it to be like a 
# left join, since there are more observations from fert_avg
fert <- full_join(fert_avg, fert_quant_avg)
fert
```


# LSMS pesticide/herbicide/fungicide data

Note: no data on quantity used is available.

```{r}
w1_phf <- sect4_pp_w1 %>% select( household_id, region=saq01, zone=saq02, prevention=pp_s4q04, pest=pp_s4q05, herb=pp_s4q06, fung=pp_s4q07) %>% drop_na(household_id, prevention) 

w2_phf <- sect4_pp_w2 %>% select(household_id2, region=saq01, zone=saq02, prevention=pp_s4q04, pest=pp_s4q05, herb=pp_s4q06, fung=pp_s4q07) %>% drop_na(household_id2, prevention) 

w3_phf <- sect4_pp_w3 %>% select(household_id2, region=saq01, zone=saq02, prevention=pp_s4q04, pest=pp_s4q05, herb=pp_s4q06, fung=pp_s4q07) %>% drop_na(household_id2, prevention) 


wave1 <- inner_join(w1_phf, loc1) 
wave1

wave2 <- inner_join(w2_phf, loc2) %>% rename(household_id = household_id2)
wave2

wave3 <- inner_join(w3_phf, loc3) %>% rename(household_id = household_id2)
wave3
```

```{r}
w1_phf <- data.frame(country="ethiopia", start_yr = 2011, end_yr = 2012, wave1, stringsAsFactors = F) 
w2_phf <- data.frame(country="ethiopia", start_yr = 2013, end_yr = 2014, wave2, stringsAsFactors = F)
phf <- data.frame(country="ethiopia", start_yr = 2015, end_yr = 2016, wave3, stringsAsFactors = F)
phf <- rbind(w1_phf, w2_phf, phf)
phf
```

```{r}
# handle missing long/lat data - only 4 cases
phf <- phf[complete.cases(phf$longitude, phf$latitude), ]
# spatial join with location data - assigns each lon/lat pair to a zone 
phf_sf <- st_as_sf(x=phf, coords=c("longitude", "latitude"), crs=4326)
phf_sf <- st_transform(phf_sf, crs=3857)

phf_joined <- st_join(phf_sf, eth_zone_sf, left=FALSE)
phf_df <- phf_joined %>% st_set_geometry(NULL) 
phf_df <- phf_df %>% mutate(NAME_1=as.character(NAME_1), NAME_2=as.character(NAME_2)) %>% select(country, start_yr, end_yr, NAME_1, NAME_2, household_id, prevention, pest, herb, fung) 
phf_df
```




```{r}
# 1 encoded as YES
# 2 encoded as NO - wish to change to 0
recode <- function(x) ifelse(x == 2, 0, 1)
phf_df <- phf_df %>% mutate_at(c("prevention", "pest", "herb", "fung"), recode)

# handling NAs - assume that if prevention is 0 then set other columns to 0
handleNA <- function(x, y) ifelse(x == 0, 0, y)
phf <- phf_df %>% mutate_at(c("pest", "herb", "fung"), list(~handleNA(prevention, .)) )


phf_df <- phf_df %>% group_by(country, start_yr, end_yr, NAME_1, NAME_2, household_id) %>% 
  summarize_all(indicator) %>% drop_na()

# compute zone avgs
phf_avg <- phf_df %>% select(-household_id) %>% group_by(country, start_yr, end_yr, NAME_1, NAME_2) %>% summarize_all(mean)

# compute zone counts
phf_avg$num_hh_phf <- (phf_df %>% tally())$n
phf_avg
```


```{r}
eth_lsms <- full_join(fert, phf_avg)
eth_lsms
```

# GAHI prevalence data
```{r}
africa <- read_csv("../GAHI/data/africa_data.csv")
eth <- africa %>% filter(country_title == "Ethiopia") %>% select(publication_year, prevalence, number_examined, number_positive, gahi_lon=longitude, gahi_lat=latitude) %>% drop_na()
eth_sf <- st_as_sf(x=eth, coords=c("gahi_lon", "gahi_lat"), crs=4326) # assume that CRS is 4326 (typical lon/lat)
eth_sf <- st_transform(eth_sf, crs=3857) # convert to 3857 (mercator/planar)
eth_sf

#plot(st_geometry(sp), axes=TRUE)
#plot(africa_pt, pch=1, col="red", add=TRUE)
joined_gahi <- st_join(eth_zone_sf, eth_sf, left=FALSE)
joined_gahi
```

```{r}
joined_gahi
```

```{r}
eth_gahi <- joined_gahi %>% select(publication_year, number_examined, number_positive, NAME_1, NAME_2) %>% group_by(publication_year, NAME_1, NAME_2) %>% summarize(prevalence=sum(number_positive)/sum(number_examined), num_examined=sum(number_examined), num_positive=sum(number_positive))
eth_gahi 
```



```{r}
eth_gahi_lsms<- inner_join(eth_gahi, eth_lsms, by = c("publication_year" = "end_yr", "NAME_1" = "NAME_1", "NAME_2" = 'NAME_2')) 

eth_gahi_lsms
```




# Final dataset 
To summarize, lsms data was outer joined together, and then lsms + gahi was inner joined.
The following rows represent entries where both GAHI and LSMS data exist. Note that we are primarily limited by the lack of GAHI data (with only one case study from 2016).

```{r}
<<<<<<< HEAD
eth_final <- eth_gahi_lsms %>% select(country, level1_name = NAME_1, level2_name=NAME_2, gahi_lsms_end_yr=publication_year, prevalence, num_positive, num_examined, fert, urea, dap, manure, compost, organic_fert, organic_fert_dummy, inorganic_fert_dummy, fert_dummy, num_hh_fert, urea_use, dap_use, iofert_use, num_hh_quant, prevention, pest, herb, fung, num_hh_phf, geometry) 
=======
eth_final <- eth_gahi_lsms %>% select(country, level2_name=NAME_2, gahi_lsms_start_yr=publication_year, lsms_end_yr=end_yr, prevalence, num_positive, num_examined, fert, urea, dap, manure, compost, organic_fert, organic_fert_dummy, inorganic_fert_dummy, fert_dummy, num_hh_fert, total_urea, total_dap, total_area, urea_use, dap_use, num_hh_quant, prevention, pest, herb, fung, num_hh_phf)
>>>>>>> 5834bf48ebdfd48e93e709424c135455df68318c
eth_final
```

REMARK: need to look at urea use for sidama, end_yr = 2014. there's an outlier there skewing the data.


# Exporting Final Data
```{r, eval=FALSE}
eth_final <- eth_final %>% st_set_geometry(NULL)
write_csv(eth_final, "../results/eth_final.csv")
```





















# Plots of GAHI+LSMS combined data

```{r}
ggplot(eth_final) + geom_sf(aes(fill = prevalence)) + facet_wrap(~gahi_lsms_end_yr) + scale_x_continuous(breaks = seq(35, 45, 4)) + ggtitle('ETH Prevalence')
```

```{r}
ggplot(eth_final) + geom_sf(aes(fill=inorganic_fert_dummy)) + facet_wrap(~gahi_lsms_end_yr) + scale_x_continuous(breaks = seq(35, 45, 4)) + ggtitle('ETH % Households using Inorganic Fertilizer')
```



```{r}
ggplot(eth_final) + geom_point(aes(x=inorganic_fert_dummy, y=prevalence)) + ggtitle('Prevalence vs Inorganic Fert')
```


# Plots of ALL data

```{r}
ggplot((eth_gahi %>% filter(publication_year >= 2011))) + geom_sf(aes(fill=prevalence)) + facet_wrap(~publication_year) + scale_x_continuous(breaks = seq(35, 45, 4)) + ggtitle('ETH GAHI Prevalence from 2011-2016')
```


```{r}
ggplot((eth_sf %>% filter(publication_year >= 2011))) + geom_sf() + facet_wrap(~publication_year) + scale_x_continuous(breaks = seq(35, 45, 4)) + ggtitle('ETH GAHI Original Locations')
```



```{r}
eth_lsms_geom <- inner_join(eth_lsms, eth_zone_sf)
g <- ggplot(eth_lsms_geom) + geom_sf(aes(fill=inorganic_fert_dummy)) + facet_wrap(~end_yr) 
g + theme(axis.text = element_blank(), axis.ticks = element_blank()) + ggtitle('ETH LSMS % Households using Inorganic Fert')
``` 


