---
title: "nigeria_lsms"
author: Elliot Quan
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE)
```

```{r}
library(sf)
library(readr)
library(dplyr)
library(tidyr)
library(dummies)
library(haven)
```



```{r}
# auxiliary function for handling sums of NA vectors
# handles NA by returning NA if all are NA
sumNA <- function(x) {
  if (all(is.na(x))) {
    return(NA)
  } else {
    return(sum(x, na.rm=TRUE))
  }
}

# take a vector of inputs: if at least 1 input is 1, then return 1 else 0.
# handles NA by returning NA if all are NA
indicator <- function(x) {
  if (all(is.na(x))) {
    return(NA)
  } else if (sum(x, na.rm=TRUE) > 0) {
    return(1)
  } else {
    return(0)
  }
}

#' Converts a column of quantity data to kg based on unit coding provided
#' by units. 1 = kg, 2 = g, 3 = liter, 4 = centiliter. Does not mutate original df.
#'
#' @param quant quantity (character) 
#' @param units specifies units of quant (integer)
#' @param data (dataframe)
#'
#' @return converted quant vector
unit_conv <- function(quant, units) {
  return(case_when(
    units == 1 ~ quant,
    units == 2 ~ quant/1000,
    units == 3 ~ quant,
    units == 4 ~ quant/100
  ))
}

#' Same as above function, except coding 1 = liter, 2 = gram, 3 = kg, other = 0 (cannot convert units)
#' This also works for 1 = kg, 2 = gram, 3 = liter
unit_conv_pest <- function(quant, units) {
  return(case_when(
    units == 1 ~ quant,
    units == 2 ~ quant/1000,
    units == 3 ~ quant,
    is.numeric(units) ~ 0 
  ))
}

#' If x is 2 or 0, set to 0. This recodes an x column.
recode <- function(x) ifelse(x == 2 | x == 0, 0, 1)

#' this is mainly used for taking a y/n column x, and wanting to return 0 if x is 0 else y for each observation.
#' for example: if x = (0, 0, 1) and y = (NA, NA, 15), then we want to return (0, 0, 15).
handleNA <- function(x, y) ifelse(x == 0, 0, y)

```

```{r}
# test 
test <- matrix(c(NA, NA, 2, 2, 3, 3, 4, 4, 5, NA), byrow=TRUE, ncol=2, nrow=5)
testdf <- data.frame(test)
testdf
testdf$'3' <- unit_conv(testdf$X1, testdf$X2)
testdf
```

# Plan

1) spatial_join lsms fert, lsms pest, and gahi prevalence data with level2 shapefiles
2) apply transformations to fert + pest to aggregate % households per level2 region
3) combine gahi prevalence with lsms data

# Remarks
Quantity data for fertilizer was NOT found (units were missing)
Wave 2 Nigeria pesticide data was missing. 
Quantity data for pesticide + herbicide can be found, but it is provided in 3 variables: amount purchased, amount leftover, and amount free.

# Shapefiles 

from GADM. data is at the 2nd administrative level (Local Govt Areas)

```{r}
nga_sf <- readRDS('../shapefiles/nga/GADM/gadm36_NGA_2_sf.rds')
nga_sf <- st_transform(nga_sf, crs=3857)
```

# Geovars (location)

```{r}
loc1 <- read_csv("../lsms_data/nigeria/w1/NGA_HouseholdGeovariables_Y1.csv")
loc1

loc2 <- read_csv("../lsms_data/nigeria/w2/NGA_HouseholdGeovars_Y2.csv")
loc2 

loc3 <- read_csv("../lsms_data/nigeria/w3/nga_householdgeovars_y3.csv")
loc3

```

```{r}
loc1_f <- loc1 %>% select(hhid, longitude=lon_dd_mod, latitude = lat_dd_mod)
loc1_mappings <- unique(loc1_f)

loc2_f <- loc2 %>% select(hhid, longitude=lon_dd_mod, latitude = lat_dd_mod)
loc2_mappings <- unique(loc2_f)

loc3_f <- loc3 %>% select(hhid, longitude=LON_DD_MOD, latitude = LAT_DD_MOD)
loc3_mappings <- unique(loc3_f)
```


# Fertilizer data

fertilizer data
```{r}
fert_w1 <- read_csv('../lsms_data/nigeria/w1/sect11d_plantingw1.csv')
fert_w2 <- read_csv('../lsms_data/nigeria/w2/sect11d_plantingw2.csv')
fert_w3 <- read_csv('../lsms_data/nigeria/w3/secta11d_harvestw3.csv')
```

area data - need to join by plotid.


```{r}
area_w1 <- read_csv('../lsms_data/nigeria/w1/sect11a1_plantingw1.csv')
area_w2 <- read_csv('../lsms_data/nigeria/w2/sect11a1_plantingw2.csv')
area_w3 <- read_csv('../lsms_data/nigeria/w3/sect11a1_plantingw3.csv')

area_w1 <- area_w1 %>% select(hhid, plotid, area = s11aq4d)
area_w2 <- area_w2 %>% select(hhid, plotid, area = s11aq4c)
area_w3 <- area_w3 %>% select(hhid, plotid, area = s11aq4c)

fert_w1 <- fert_w1 %>% inner_join(area_w1, by = c('hhid', 'plotid'))
fert_w2 <- fert_w2 %>% inner_join(area_w2, by = c('hhid', 'plotid'))
fert_w3 <- fert_w3 %>% inner_join(area_w3, by = c('hhid', 'plotid'))
```




```{r}
fert1 <- fert_w1 %>% select(zone, state, hhid, plotid, area, fert_yn = s11dq1, 
                          leftover_yn = s11dq2, leftover_type = s11dq3, leftover_quant = s11dq4, 
                          free_yn = s11dq6, free_type = s11dq7, free_quant = s11dq8,
                          bought1_yn = s11dq12, bought1_type = s11dq14, bought1_quant = s11dq15,
                          bought2_yn = s11dq23, bought2_type = s11dq25, bought2_quant = s11dq26) %>% drop_na(fert_yn)


fert2 <- fert_w2 %>% select(zone, state, hhid, plotid, area, fert_yn = s11dq1, 
                            leftover_yn = s11dq2, leftover_type = s11dq3, leftover_quant = s11dq4,
                            free_yn = s11dq6, free_type = s11dq7, free_quant = s11dq8,
                            bought1_yn = s11dq12, bought1_type = s11dq15, bought1_quant = s11dq16,
                            bought2_yn = s11dq24, bought2_type = s11dq27, bought2_quant = s11dq28) %>% drop_na(fert_yn)

# wave3 quantities are varied, so need to standardize all weights to kg using unit_conv
fert3 <- fert_w3 %>% select(zone, state, hhid, plotid, area, fert_yn = s11dq1, 
                            leftover_yn = s11dq2, leftover_type = s11dq3, leftover_quant = s11dq4a, leftover_quant_units = s11dq4b,
                            free_yn = sect11dq6, free_type = sect11dq7, free_quant = sect11dq8a, free_quant_units = sect11dq8b,
                            bought1_yn = s11dq12, bought1_type = s11dq15, bought1_quant = s11dq16a, bought1_quant_units = s11dq16b,
                            bought2_yn = s11dq24, bought2_type = s11dq27, bought2_quant = s11dq28a, bought2_quant_units = s11dq28b) %>% drop_na(fert_yn) %>% mutate(leftover_quant = unit_conv(leftover_quant, leftover_quant_units),
                          free_quant = unit_conv(free_quant, free_quant_units),
                          bought1_quant = unit_conv(bought1_quant, bought1_quant_units), 
                          bought2_quant = unit_conv(bought2_quant, bought2_quant_units)) %>% select(-contains('units'))

fert1_joined <- inner_join(fert1, loc1_mappings)
fert1_joined

fert2_joined <- inner_join(fert2, loc2_mappings)
fert2_joined

fert3_joined <- inner_join(fert3, loc3_mappings)
fert3_joined
```

```{r}
w1_fert <- data.frame(country="nigeria", start_yr = 2010, end_yr = 2011, fert1_joined, stringsAsFactors = F) 
w2_fert <- data.frame(country="nigeria", start_yr = 2012, end_yr = 2013, fert2_joined, stringsAsFactors = F)
w3_fert <- data.frame(country="nigeria", start_yr = 2015, end_yr = 2016, fert3_joined, stringsAsFactors = F)
fert <- rbind(w1_fert, w2_fert, w3_fert)
fert[!complete.cases(fert$longitude, fert$latitude), ]
fert
```

# aggregating all quantities
For each of leftover, free, bought1, bought2:
  1) want to calculate what type of fertilizer was used
  2) add that quantity to the corresponding type total
  
There are 4 types: npk, urea, composite manure, other.


if no fert_yn == 2, then want to set other yn cols to 0.

```{r}
# for all y/n cols, change 2s to 0s

fert <- fert %>% mutate_at(c('fert_yn', 'leftover_yn', 'free_yn', 'bought1_yn', 'bought2_yn'), recode)

# if no fert used, then set the following cols to 0.

fert <- fert %>% mutate_at(c('leftover_yn', 'free_yn', 'bought1_yn', 'bought2_yn', 'leftover_type', 'free_type', 'bought1_type', 'bought2_type', 'leftover_quant', 'free_quant', 'bought1_quant', 'bought2_quant'), list(~handleNA(fert_yn, .)))


# same goes for the other categories of use: if none used, set both type and quant to 0.
fert <- fert %>% mutate_at(c('leftover_type', 'leftover_quant'), list(~handleNA(leftover_yn, .))) %>% 
  mutate_at(c('free_type', 'free_quant'), list(~handleNA(free_yn, .))) %>%
  mutate_at(c('bought1_type', 'bought1_quant'), list(~handleNA(bought1_yn, .))) %>%
  mutate_at(c('bought2_type', 'bought2_quant'), list(~handleNA(bought2_yn, .)))
fert
```

```{r}
# for each of leftover, free, bought1, bought2, check that the type matches the desired type. if so, then add that quantity to the sum and return the sum.
agg_quant <- function(type, df) {
  return(ifelse(df$leftover_type == type, df$leftover_quant, 0) + ifelse(df$free_type == type, df$free_quant, 0) + ifelse(df$bought1_type == type, df$bought1_quant, 0) + ifelse(df$bought2_type == type, df$bought2_quant, 0))
}

# return 1 if at least one of the 4 columns equals type, else 0.
yn_quant <- function(type, df) {
  return(as.numeric(((df$leftover_type == type) | (df$free_type == type) | (df$bought1_type == type) | (df$bought2_type == type))))
}


fert <- fert %>% mutate(npk_yn  = yn_quant(1, fert), urea_yn = yn_quant(2, fert), comp_manure_yn = yn_quant(3, fert), other_yn = yn_quant(4, fert))

fert <- fert %>% mutate(npk_quant = agg_quant(1, fert), urea_quant = agg_quant(2, fert), comp_manure_quant = agg_quant(3, fert), other_quant = agg_quant(4, fert), total_quant = npk_quant + urea_quant + comp_manure_quant + other_quant)

fert
```
  
```{r}
fert <- fert %>% select(country, start_yr, end_yr, hhid, plotid, area, fert_yn, npk_yn, urea_yn, comp_manure_yn, other_yn, npk_quant, urea_quant, comp_manure_quant, other_quant, total_quant, longitude, latitude)
fert
```



```{r}
# join with sf data

fert_sf <- st_as_sf(x=fert, coords=c("longitude", "latitude"), crs = 4326)
fert_sf <- st_transform(fert_sf, crs= 3857)
fert_joined <- st_join(nga_sf, fert_sf, left=FALSE)
fert_joined
```


```{r}
fert_df <- fert_joined %>% select(country, start_yr, end_yr, NAME_1, NAME_2, hhid, area, fert_yn, npk_yn, urea_yn, comp_manure_yn, other_yn, npk_quant, urea_quant, comp_manure_quant, other_quant, total_quant) %>% st_set_geometry(NULL)
rownames(fert_df) <- 1:nrow(fert_df)

fert_df_use <- fert_df %>% group_by(country, start_yr, end_yr, NAME_1, NAME_2, hhid) %>% summarize(npk_use = mean(npk_quant/area), urea_use = mean(urea_quant/area), comp_manure_use = mean(comp_manure_quant/area), other_use = mean(other_quant/area), total_use = mean(total_quant/area)) %>% drop_na()

fert_df_indic  <- fert_df %>% group_by(country, start_yr, end_yr, NAME_1, NAME_2, hhid) %>% summarize_each(funs=(indicator), fert_yn, npk_yn, urea_yn, comp_manure_yn, other_yn) %>% drop_na()

fert_df
fert_df_use
fert_df_indic
```

```{r}
# creating indicator vars for organic, inorganic, and any fert
fert_df_indic <- fert_df_indic %>% mutate(organic_fert_dummy = comp_manure_yn, inorganic_fert_dummy = as.numeric((npk_yn + urea_yn) > 0), fert_dummy = as.numeric((organic_fert_dummy + inorganic_fert_dummy + other_yn) > 0)) 

# grouping by hh: indicators. also dropping NAs at this point
fert_only <- fert_df_indic %>% select(-hhid) %>% group_by(country, start_yr, end_yr, NAME_1, NAME_2) %>% summarize_all(mean)
fert_only$num_hh_fert <- (fert_df_indic %>% tally())$n

# grouping by hh: quantities. also dropping NAs at this point.
fert_quant <- fert_df_use %>% select(-hhid) %>% group_by(country, start_yr, end_yr, NAME_1, NAME_2) %>% summarize_all(mean)
fert_quant$num_hh_quant <- (fert_df_use %>% tally())$n

fert_only
fert_quant
```

```{r}
fert_all <- inner_join(fert_only, fert_quant)
fert_all
```

```{r}
for (i in 1:ncol(fert_all)) {
  print(sum(is.na(fert_all[, i])))
}

```

# Pesticide/Herbicide

Note: wave 2 is missing sect11c2_plantingw2 -  will try to redownload.

```{r}
w1_pest <- read_csv('../lsms_data/nigeria/w1/sect11c_plantingw1.csv')
w1_pest

w3_pest <- read_csv('../lsms_data/nigeria/w3/secta11c2_harvestw3.csv')
w3_pest
```


```{r}
pest1 <- w1_pest %>% select(hhid, plotid, pest_yn = s11cq1, pest_quant = s11cq2a, pest_quant_unit = s11cq2b, herb_yn = s11cq10, herb_quant = s11cq11a, herb_quant_unit = s11cq11b) %>% drop_na(pest_yn, herb_yn)
pest3 <- w3_pest %>% select(hhid, plotid, pest_yn =s11c2q1, pest_quant = s11c2q2a, pest_quant_unit = s11c2q2b, herb_yn = s11c2q10, herb_quant = s11c2q11a, herb_quant_unit = s11c2q11b) %>% drop_na(pest_yn, herb_yn)
pest1
pest3

pest1_j <- inner_join(pest1, loc1_mappings)
pest3_j <- inner_join(pest3, loc3_mappings)
pest1_j
pest3_j
```



```{r}
pest_w1 <- pest1_j %>% inner_join(area_w1, by = c('hhid', 'plotid'))

pest_w3 <- pest3_j %>% inner_join(area_w3, by = c('hhid', 'plotid'))

pest_w1
pest_w3
```

```{r}
w1_p <- data.frame(country="nigeria", start_yr = 2010, end_yr = 2011, pest_w1, stringsAsFactors = F) 
w3_p <- data.frame(country="nigeria", start_yr = 2015, end_yr = 2016, pest_w3, stringsAsFactors = F)
pest_c <- rbind(w1_p, w3_p)
pest_c
```


```{r}
# for y/n cols, set 2 to 0
pest_c[pest_c$pest_yn == 2, c('pest_yn')] <- 0
pest_c[pest_c$herb_yn == 2, c('herb_yn')] <- 0

# if no pest used or no herb used, then set quant and type columns to 0.
pest_c <- pest_c %>% mutate_at(c('pest_quant', 'pest_quant_unit'), list(~handleNA(pest_yn, .))) %>% 
  mutate_at(c('herb_quant', 'herb_quant_unit'), list(~handleNA(herb_yn, .)))
pest_c
```

```{r}
pest_conv <- pest_c %>% mutate(pest_quant = unit_conv_pest(pest_quant, pest_quant_unit), herb_quant = unit_conv_pest(herb_quant, herb_quant_unit))
pest_conv
```

```{r}
pest_conv[!complete.cases(pest_conv), ]
```



```{r}
pest_sf <- st_as_sf(x=pest_conv, coords=c("longitude", "latitude"), crs = 4326)
pest_sf <- st_transform(pest_sf, crs=3857)
pest_joined <- st_join(nga_sf, pest_sf, left=FALSE)
pest_joined
```

```{r}
pest_df <- pest_joined %>% select(country, start_yr, end_yr, NAME_1, NAME_2, hhid, pest_yn, pest_quant, herb_yn, herb_quant, area) %>% st_set_geometry(NULL)
rownames(pest_df) <- 1:nrow(pest_df)
pest_df
```

```{r}
# separate into indicators (Y/N) vs quantity data. 
# for indicators, we first aggregate rows to be households : 1 if household used, else 0.
pest_indic <- pest_df %>% select(-c('pest_quant', 'herb_quant', 'area')) %>% group_by(country, start_yr, end_yr, NAME_1, NAME_2, hhid) %>% summarize_all(indicator)


# for quantity, we aggregate rows to be households, but we just take the mean use (quant/area)
pest_quant <- pest_df %>% select(-c('pest_yn', 'herb_yn')) %>% group_by(country, start_yr, end_yr, NAME_1, NAME_2, hhid) %>% summarize(pest_use = mean(pest_quant/area), herb_use = mean(herb_quant/area)) 


# now we can group by NAME_2 and compute an average over households
pest_indic_grouped <- pest_indic %>% select(-hhid) %>% group_by(country, start_yr, end_yr, NAME_1, NAME_2) %>% summarize_all(mean) 
pest_indic_grouped$num_hh_indic <- (pest_indic %>% tally())$n

# same idea
pest_quant_grouped <- pest_quant %>% select(-hhid) %>% group_by(country, start_yr, end_yr, NAME_1, NAME_2) %>% summarize_all(mean)
pest_quant_grouped$num_hh_quant <- (pest_quant %>% tally())$n


pest_all <- pest_indic_grouped %>% inner_join(pest_quant_grouped, by = c('end_yr', 'NAME_2')) %>% select(-c('country.y', 'start_yr.y', 'NAME_1.y')) %>% rename(country = country.x, start_yr = start_yr.x, NAME_1 = NAME_1.x)
pest_all
```



```{r}
nga_lsms <- full_join(fert_all, pest_all)
nga_lsms
```



# GAHI prevalence

```{r}
africa <- read_csv("../GAHI/data/africa_data.csv")
nga <- africa %>% filter(country_title == "Nigeria") %>% select(publication_year, prevalence, number_examined, number_positive, longitude, latitude) %>% drop_na()
nga_gahi_sf <- st_as_sf(x=nga, coords=c("longitude", "latitude"), crs= 4326)
nga_gahi_sf <- st_transform(nga_gahi_sf, crs = 3857)
#plot(st_geometry(sp), axes=TRUE)
#plot(africa_pt, pch=1, col="red", add=TRUE)
joined_gahi <- st_join(nga_sf, nga_gahi_sf, left=FALSE)
joined_gahi
```



```{r}
# recalculate average prevalence per level2 
nga_gahi <- joined_gahi %>% select(publication_year, number_examined, number_positive, NAME_1, NAME_2) %>% group_by(publication_year, NAME_1, NAME_2) %>% summarize(prevalence=sum(number_positive)/sum(number_examined), num_examined=sum(number_examined), num_positive=sum(number_positive))
nga_gahi 
```

```{r}
nga_lsms[nga_lsms$end_yr > 2010, ]
```


```{r}
nga_gahi_lsms <- inner_join(nga_gahi, nga_lsms, by=c('publication_year' = 'end_yr', 'NAME_1', 'NAME_2')) %>% st_set_geometry(NULL)
nga_gahi_lsms
```

```{r}
nga_final <- nga_gahi_lsms %>% select(country, level2_name=NAME_2, gahi_lsms_start_yr=publication_year, lsms_end_yr=end_yr, prevalence, num_positive, num_examined, fert, npk, urea, composite_manure, other, organic_fert_dummy, inorganic_fert_dummy, fert_dummy, num_hh_fert, pest, herb, num_hh_phf)
nga_final
```

```{r, eval=FALSE}
write_csv(nga_final, "../results/nga_final.csv")
```



# DTA to CSV
```{r, eval=FALSE}
a <- read_dta('../lsms_data/nigeria/dta/sect11d_plantingw1.dta')
write_csv(a, '../lsms_data/nigeria/sect11d_plantingw1.csv')

a <- read_dta('../lsms_data/nigeria/dta/sect11d_plantingw2.dta')
write_csv(a, '../lsms_data/nigeria/sect11d_plantingw2.csv')

a <- read_dta('../lsms_data/nigeria/dta/sect11c_plantingw1.dta')
write_csv(a, '../lsms_data/nigeria/sect11c_plantingw1.csv')

a <- read_dta('../lsms_data/nigeria/dta/sect11c1_plantingw2.dta')
write_csv(a, '../lsms_data/nigeria/sect11c1_plantingw2.csv')
```

